Hydra Battles and Cie

### 2021-08-24 doc/part-hydras.tex
Alectryon: 75% of chapter6

\begin{Coqsrc}
Definition L_omega_square k := iterate (fun z => S (2 * z)%nat)
                                        k
                                        (S k).

Lemma L_omega_square_eqn k :
  (0 < k)%nat ->
  L_omega_square k = (exp2 k * (k + 2) - 1)%nat.

Lemma L_omega_square_ok: L_spec (T1.omega * T1.omega) 
          L_omega_square.

Compute L_omega_square 8.
\end{Coqsrc}

\begin{Coqanswer}
 = 2559
     : nat
\end{Coqanswer}

@@

\begin{Coqsrc}
Section phi0_mult.
 Variables (alpha : T1) (f : nat -> nat).
 Hypotheses (Halpha : nf alpha)
            (f_mono : strict_mono f)
            (f_Sle : S <<= f)
            (f_ok : L_spec (phi0 alpha) f).

 Definition L_phi0_mult i := iterate f i.

Lemma L_phi0_mult_ok i: 
  L_spec (ocons alpha i zero)  (L_phi0_mult (S i)).

 Lemma L_phi0_mult_smono i: strict_mono (L_phi0_mult i).

 Lemma L_phi0_mult_Sle i: S <<=  L_phi0_mult (S i).

End phi0_mult.
\end{Coqsrc}

Let us look 
at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}

\begin{Coqsrc}
 Definition L_omega_square_times i :=  iterate L_omega_square i.

 Lemma L_omega_square_times_ok i : 
    L_spec (ocons 2 i zero) (L_omega_square_times (S i)).
 Proof.
  apply L_phi0_mult_ok.
  -  auto with T1.
  -  apply L_omega_square_Sle.
  -  apply L_omega_square_ok.
 Qed.
\end{Coqsrc}

We are now ready to get an exact formula for $L_{\omega^3}$. 
\begin{Coqsrc}
Definition L_omega_cube  := L_lim  L_omega_square_times .

Lemma L_omega_cube_ok : L_spec (phi0 3) L_omega_cube.
\end{Coqsrc}

The function  $L_{\omega^3}$  is just obtained by diagonalization upon $L_{\omega^2\times i}$.

\begin{Coqsrc}
Lemma L_omega_cube_eqn i : 
   L_omega_cube i = L_omega_square_times i (S i).
Proof. reflexivity. Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_omega_cube_3_eq:
   let N := exp2 95 in
   let P := (N * 97 - 1)%nat in
   L_omega_cube 3  =  (exp2 P * (P + 2) - 1)%nat.
\end{Coqsrc}

@@

\label{Functions:L-alpha}
 \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}
\begin{Coqsrc}
From Equations Require Import Equations.
Require Import ArithRing Lia.

Instance Olt : WellFounded Lt := Lt_wf.
Hint Resolve Olt : E0.

(** Using Coq-Equations for building a function which satisfies 
    Large_sets.L_spec *)

Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  L_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  L_ (Canon alpha i)  (S i) ;
            | right notlimit =>  L_ (Pred alpha) (S i)}}.

It is worth looking at the answer from \texttt{Equations} and at all the lemmas this plug-in gives you for free. We show here only a part of \coq's answer.

\begin{Coqanswer}
L__obligations_obligation_1 is defined
L__obligations_obligation_2 is defined
L__obligations is defined
L__clause_1 is defined
L__functional is defined
L_ is defined
...
L__equation_1 is defined
L__graph_mut is defined
L__graph_rect is recursively defined
L__graph_correct is defined
L__elim has type-checked, generating 1 obligation
L__elim is defined
FunctionalElimination_L_ is defined
FunctionalInduction_L_ is defined
\end{Coqanswer}

Sometimes, these automatically generated statements may look cryptic.

\begin{Coqsrc}
About L__equation_1.
\end{Coqsrc}

\begin{Coqanswer}
L__equation_1 :
forall (alpha : E0) (i : nat),
L_ alpha i = L__unfold_clause_1 alpha (E0_eq_dec alpha Zero) i
\end{Coqanswer}

@@

\begin{Coqsrc}
Lemma L_zero_eqn : forall i, L_ Zero i = i.
Proof. intro i; now rewrite L__equation_1. Qed.

Lemma L_lim_eqn alpha i : Limitb alpha -> L_ alpha i =
                                        L_ (Canon alpha i) (S i).

Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).

Hint Rewrite L_zero_eqn L_succ_eqn : L_rw.
\end{Coqsrc}

>>>>

\input{movies/snippets/Large_Sets/LOmegaSquare}
\input{movies/snippets/Large_Sets/LOmegaSquareEqn}
\input{movies/snippets/Large_Sets/LOmegaSquareOk}

@@

\input{movies/snippets/Large_Sets/phi0Mult}

Let us look now
at the ordinal $\omega^2\times i$, using \texttt{L\_phi0\_mult}.

\input{movies/snippets/Large_Sets/LOmegaSquareTimes}

We are now ready to get an exact formula for $L_{\omega^3}$, by diagonalization upon $L_{\omega^2\times i}$.

\input{movies/snippets/Large_Sets/LOmegaCube}

@@

\input{movies/snippets/Large_Sets/LOmegaCube3Eq}

@@

\label{Functions:L-alpha}
 \index{hydras}{Library Epsilon0!Functions!L\_@L\_ (final step of a minimal path}

\input{movies/snippets/L_alpha/LDef}

This definition results in a bunch of automatically generated lemmas. For instance:

\input{movies/snippets/L_alpha/AboutLEquation1}

@@

\input{movies/snippets/L_alpha/Paraphrases}
